### 2023-01-18 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@e7b136a5043c607ef450d8b87226b0e4e956ead9 🚀

In this section, we demonstate how differences in species'
relative-abundance can lead to many false-negatives in interaction data.
We also introduce a method for producing a null estimate of the
false-negative-rate in datasets via simulation. Because the true
false-negative-rate is latent, we can never actually be sure how many
false-negatives are in our data. However, here we outline an approach to
deal with this fact---first by using simulation to estimate the
false-negative-rate for a dataset of a fixed size using neutral models
of observation. We then illustrate how to incorporate uncertainty
directly into predictions of species interactions to account for
observation error based on null estimates of both false-positive rates
(as an priori estimate of species misidentification probability) and
false-positive rates (as generated via the method we introduce).

@@

We start with a naive model of interaction detection: we assume that
every interacting pair of species is incorrectly observed as
not-interacting with an independent and fixed probability, which we
denote \(p_{fn}\) and subsequently refer to as the False-Negative-Rate
(FNR). If we observe the same species not-interacting \(N\) times, then
the probability of a true-negative (denoted \(p_{tn}\)) is given by
\(p_{tn}=1-(p_{fn})^N\). This relation (the probability-mass-function of
geometric distribution, a special case of the negative-binomial
distribution) is shown in fig.~\ref{fig:geometric}(a) for varying values
of \(p_{fn}\) and illustrates a fundamental link between our ability to
reliably say an interaction doesn't exist---\(p_{tn}\)---and the number
of times \(N\) we have observed a given species. In addition, note that
there is no non-zero \(p_{fn}\) for which we can ever \emph{prove} that
an interaction does not exist---no matter how many observations of
non-interactions \(N\) we have, \(p_{tn}<1\).

From fig.~\ref{fig:geometric}(a) it is clear that the more often we see
two species co-occurring, but \emph{not interacting}, the more likely
the interaction is a true-negative. This has several practical
consequences: first it means negatives taken outside the overlap of the
range of each species aren't informative because co-occurrence was not
possible, and therefore neither was an interaction. If some of the
``worst-case'' FNRs presented in fig.~\ref{fig:geometric}(a) seem
unrealistically high, considering that species are observed in
proportion to their relative abundance. In the next section we
demonstrate that the distribution of abundance in ecosystems can lead to
very high realized values of FNR (\(p_{fn}\)) simply as an artifact of
sampling effort. Second, we can use this relation to compute the
expected number of total observations needed to obtain a ``goal'' number
of observations of a particular pair of species
(fig.~\ref{fig:geometric}(b)). As an example, if we hypothesize that
\(A\) and \(B\) do not interact, and we want to see species \(A\) and
\(B\) both co-occurring and \emph{not interacting} 10 times to be
confident this is a true negative, then we need an expected 1000
observations of all species if the relative abundances of \(A\) and
\(B\) are both \(0.1\).

@@

We now show that the realized FNR changes drastically with sampling
effort due to the intrinsic variation of the abundance of individuals of
each species within a community. We do this by simulating the process of
observation of species interactions, applied both to 243 empirical food
webs from the Mangal database (Banville \emph{et al.} 2021) and random
food-webs generated using the niche model, a simple generative model of
food-web structure that accounts for allometric scaling (Williams \&
Martinez 2000). Our neutral model of observation assumes each observed
species is drawn in proportion to each species' abundance at that place
and time. The abundance distribution of a community can be
reasonably-well described by a log-normal distribution (Volkov \emph{et
al.} 2003). In addition to the log-normal distribution, we also tested
the case where the abundance distribution is derived from power-law
scaling \(Z^{(log(T_i)-1)}\) where \(T_i\) is the trophic level of
species \(i\) and \(Z\) is a scaling coefficient (Savage \emph{et al.}
2004), which yields the same qualitative behavior. The practical
consequence of abundance distributions spanning many orders of magnitude
of abundance is that observing two ``rare'' species interacting requires
two low probability events: observing two rare species \emph{at the same
time}.

To simulate the process of observation, for an ecological network \(M\)
with \(S\) species, we sample abundances for each species from a
standard-log-normal distribution. For each true interaction in the
adjacency matrix \(M\) (i.e. \(M_{ij}=1\)) we estimate the probability
of observing both species \(i\) and \(j\) at a given place and time by
simulating \(n\) observations of all individuals of any a species, where
the species of the individual observed at the \(\{1,2,\dots,n\}\)-th
observation is drawn from the generated log-normal distribution of
abundances. For each pair of species \((i,j)\), if both \(i\) and \(j\)
are observed within the n-observations, the interaction is tallied as a
true positive if \(M_{ij}=1\). If only one of \(i\) or \(j\) are
observed---but not both---in these \(n\) observations, but \(M_{ij}=1\),
this is counted as a false-negative, and a true-negative otherwise. For
each pair of species \((i,j)\), if both \(i\) and \(j\) are observed
within the n-observations, the interaction is tallied as a true positive
if \(M_{ij}=1\). If only one of \(i\) or \(j\) are observed---but not
both---in these \(n\) observations, but \(M_{ij}=1\), this is counted as
a false-negative, and a true-negative otherwise (\(M_{ij} = 0\)). This
process is illustrated conceptually in
fig.~\ref{fig:resampling_concept}(a).

@@

where the maximum \emph{a posteriori} (MAP) estimate of \((\mu, \phi)\)
applied to Mangal data from (MacDonald \emph{et al.} 2020) is
\((\mu=0.086, \phi=24.3)\). All simulations were done with 500
independent replicates of unique niche model networks per unique number
of observations \(n\). All analyses presented here are done in Julia
v1.8 (Bezanson \emph{et al.} 2015) using both EcologicalNetworks.jl v0.5
and Mangal.jl v0.4 (Banville \emph{et al.} 2021) and are hosted on
\href{https://github.com/gottacatchenall/ms_false_negatives/tree/main/src}{Github}).
Note that the empirical data, for the reasons described above, very
likely already contains many false-negatives, we'll revisit this issue
in the final section.

From fig.~\ref{fig:geometric}(c) it is evident that the number of
species considered in a study is inseparable from the
false-negative-rate in that study, and this effect should be taken into
account when designing samples of ecological networks in the future. We
see a similar qualitative pattern in fig.~\ref{fig:geometric}(d) where
the FNR drops off quickly as a function of observation effort, mediated
by total richness. The practical consequence of the bottom row of
fig.~\ref{fig:geometric} is whether the total number of observations of
all species (the x-axis) for the threshold FNR we deem acceptable (the
y-axis) is feasible. This raises two points: first, empirical data on
interactions are subject to the practical limitations of funding and
human-work hours, and therefore existing data tend to fall on the order
of hundreds or thousands observations of individuals per site. Clear
aggregation of data on sampling effort has proven difficult to find and
a meta-analysis of network data and sampling effort seems both pertinent
and necessary, in addition to the effects of aggregation of interactions
across taxonomic scales (Gauzens \emph{et al.} 2013; Giacomuzzo \&
Jordán 2021). This inherent limitation on in-situ sampling means we
should optimize where we sample across space so that for a given number
of samples, we obtain the maximum information possible. Second, what is
meant by ``acceptable'' FNR? This raises the question: does a shifting
FNR lead to rapid transitions in our ability inference and predictions
about the structure and dynamics of networks, or does it produce a
roughly linear decay in model efficacy? We explore this in the next
section.

We conclude this section by advocating for the use of neutral models
similar to above to generate expectations about the number of
false-negatives in a data set of a given size. This could prove fruitful
both for designing surveys of interactions but also because we may want
to incorporate models of imperfect detection error into predictive
interactions models, as Joseph (2020) does for species occurrence
modeling. Additionally, we emphasize that one must consider the context
for sampling---is the goal to detect a particular species (as in
fig.~\ref{fig:geometric}(c)), or to get a representative sample of
interactions across the species pool? These arguments are
well-considered when sampling individual species (Willott 2001), but
have not yet been adopted for designing samples of communities.

@@

Here we show how to incorporate uncertainty into model predictions of
interaction probability to account for imperfect observation (both
false-negatives and false-positives). Models for interaction prediction
typically yield a probability of interaction between each pair of
species, \(p_{ij}\). When these are considered with uncertainty, it is
usually model-uncertainty, e.g.~the variance in the interaction
probability prediction across several cross-validation folds, where the
data is split into training and test sets several times. The method we
introduce adjusts the value of a model's predictions to produce a
distribution of interaction probabilities, which are adjusted by a given
false-negative-rate \(p_{fn}\) and false-positive-rate \(p_{fp}\)
(outlined in figure fig.~\ref{fig:resampling_concept}). We describe
first how to sample from this distribution of adjusted interaction
probabilities via simulation, and show that this distribution can be
well-approximated analytically.

@@

We then consider the output prediction from an arbitrary prediction
model, which is the probability \(p_{ij}\) that two species \(i\) and
\(j\) interact. To get an estimate of \(p_{ij}\) that accounts for
observation error, we resample the probability of each interaction
\(p_{ij}\) by simulating a set of several `particles,' where each
particle is a realization of an interaction occurring (either true or
false with probabilities \(p_{ij}\) and \(1-p_{ij}\) respectively) and
then being correctly observed with probabilities given by \(p_{fp}\) and
\(p_{fn}\) to yield a single boolean outcome for each particle
(``Resampling'' within fig.~\ref{fig:resampling_concept}(b)). Across of
many particles, the resulting frequency of `true' outcomes is a single
resample of the interaction probability \(p_{ij}^*\). Across several
samples each of several particles, this forms a distribution of
probabilities which are adjusted by the true and false negative rates.

@@

To test this we use the dataset from Hadfield \emph{et al.} (2014) that
describes host-parasite interaction networks sampled across 51 sites,
and the same method as Strydom \emph{et al.} (2021) to extract latent
features for each species in this dataset based on applying PCA to the
co-occurrence matrix. We then predict a metaweb (equivalent to
predicting true or false for an interaction between each species pair,
effectively a binary classification problem) from these species-level
features using four candidate models for binary classification---three
often used machine-learning (ML) methods (Boosted Regression Tree (BRT),
Random Forest (RF), Decision Tree (DT)), and one naive model from
classic statistics (Logistic Regression (LR)). Each of the ML models are
bootstrap aggregated (or bagged) with 100 replicates each. We partition
the data into 80-20 training-test split, and then seed the training data
with false negatives at varying rates, but crucially do nothing to the
test data. We fit all of these models using MLJ.jl, a high-level Julia
framework for a wide-variety of ML models (Blaom \emph{et al.} 2020). We
evaluate the efficacy of these models using two common measures of
binary classifier performance: the area under the receiver-operator
curve (ROC-AUC) and the area under the precision-recall curve (PR-AUC),
for more details see Poisot (2022). Here, PR-AUC is slightly more
relevant as it is a better indicator of prediction of false-negatives.
The results of these simulations are shown in fig.~\ref{fig:addedfnr}(a
\& b).

>>>>

In this section, we demonstate how differences in species'
relative-abundance can lead to many false-negatives in interaction data.
We also introduce a method for producing a null estimate of the
false-negative-rate in datasets via simulation. Because the true
false-negative-rate is latent, we can never actually be sure how many
false-negatives are in our data. However, here we outline an approach to
deal with this fact---first by using simulation to estimate the
false-negative-rate for a dataset of a fixed size using neutral models
of observation. We then illustrate how to incorporate uncertainty
directly into predictions of species interactions to account for
observation error based on null estimates of both the false-positive
rate (as an \emph{a priori} estimate of species misidentification
probability) and false-negative rate (as generated via the method we
introduce).

@@

We start with a naive model of interaction detection: we assume that
every interacting pair of species is incorrectly observed as
not-interacting with an independent and fixed probability, which we
denote \(p_{fn}\) and subsequently refer to as the False-Negative-Rate
(FNR). If we observe the same species not-interacting \(N\) times, then
the probability of a true-negative (denoted \(p_{tn}\)) is given by
\(p_{tn}=1-(p_{fn})^N\). This relation (the
cumalitive-distribution-function of geometric distribution, a special
case of the negative-binomial distribution) is shown in
fig.~\ref{fig:geometric}(a) for varying values of \(p_{fn}\) and
illustrates a fundamental link between our ability to reliably say an
interaction doesn't exist---\(p_{tn}\)---and the number of times \(N\)
we have observed a given species. In addition, note that there is no
non-zero \(p_{fn}\) for which we can ever \emph{prove} that an
interaction does not exist---no matter how many observations of
non-interactions \(N\) we have, \(p_{tn}<1\).

From fig.~\ref{fig:geometric}(a) it is clear that the more often we see
two species co-occurring, but \emph{not interacting}, the more likely
the interaction is a true-negative. This has several practical
consequences: first it means negatives taken outside the overlap of the
range of each species aren't informative because co-occurrence was not
possible, and therefore neither was an interaction. In the next section
we demonstrate that the distribution of abundance in ecosystems can lead
to very high realized values of FNR (\(p_{fn}\)) simply as an artifact
of sampling effort. Second, we can use this relation to compute the
expected number of total observations needed to obtain a ``goal'' number
of observations of a particular pair of species
(fig.~\ref{fig:geometric}(b)). As an example, if we hypothesize that
\(A\) and \(B\) do not interact, and we want to see species \(A\) and
\(B\) both co-occurring and \emph{not interacting} 10 times to be
confident this is a true negative, then we need an expected 1000
observations of all species if the relative abundances of \(A\) and
\(B\) are both \(0.1\).

@@

We now show that the realized FNR changes drastically with sampling
effort due to the intrinsic variation of the abundance of individuals of
each species within a community. We do this by simulating the process of
observation of species interactions, applied both to 243 empirical food
webs from the Mangal database (Banville \emph{et al.} 2021) and random
food-webs generated using the niche model, a simple generative model of
food-web structure that accounts for allometric scaling (Williams \&
Martinez 2000). Our neutral model of observation assumes each observed
species is drawn in proportion to each species' abundance at that place
and time. The abundance distribution of a community can be
reasonably-well described by a log-normal distribution (Volkov \emph{et
al.} 2003). In addition to the log-normal distribution, we also tested
the case where the abundance distribution is derived from power-law
scaling \(Z^{(log(T_i)-1)}\) where \(T_i\) is the trophic level of
species \(i\) and \(Z\) is a scaling coefficient (Savage \emph{et al.}
2004), which yields the same qualitative behavior. The practical
consequence of abundance distributions spanning many orders of magnitude
is that observing two ``rare'' species interacting requires two low
probability events: observing two rare species \emph{at the same time}.

To simulate the process of observation, for an ecological network \(M\)
with \(S\) species, we sample relative abundances for each species from
a standard-log-normal distribution. For each true interaction in the
adjacency matrix \(M\) (i.e.~\(M_{ij}=1\)) we estimate the probability
of observing both species \(i\) and \(j\) at a given place and time by
simulating \(n\) observations of all individuals of any species, where
the species of the individual observed at the \(\{1,2,\dots,n\}\)-th
observation is drawn from the generated categorical distribution of
abundances. For each pair of species \((i,j)\), if both \(i\) and \(j\)
are observed within the n-observations, the interaction is tallied as a
true positive if \(M_{ij}=1\). If only one of \(i\) or \(j\) are
observed---but not both---in these \(n\) observations, but \(M_{ij}=1\),
this is counted as a false-negative, and a true-negative otherwise. For
each pair of species \((i,j)\), if both \(i\) and \(j\) are observed
within the n-observations, the interaction is tallied as a true positive
if \(M_{ij}=1\). If only one of \(i\) or \(j\) are observed---but not
both---in these \(n\) observations, but \(M_{ij}=1\), this is counted as
a false-negative, and a true-negative otherwise (\(M_{ij} = 0\)). This
process is illustrated conceptually in
fig.~\ref{fig:resampling_concept}(a).

@@

where the maximum \emph{a posteriori} (MAP) estimate of \((\mu, \phi)\)
applied to Mangal data from (MacDonald \emph{et al.} 2020) is
\((\mu=0.086, \phi=24.3)\). All simulations were done with 500
independent replicates of unique niche model networks per unique number
of total interactions observed \(n\). All analyses presented here are
done in Julia v1.8 (Bezanson \emph{et al.} 2015) using both
EcologicalNetworks.jl v0.5 and Mangal.jl v0.4 (Banville \emph{et al.}
2021) and are hosted on
\href{https://github.com/gottacatchenall/ms_false_negatives/tree/main/src}{Github}).
Note that the empirical data, for the reasons described above, very
likely already contains many false-negatives, we'll revisit this issue
in the final section.

From fig.~\ref{fig:geometric}(c) it is evident that the number of
species considered in a study is inseparable from the
false-negative-rate in that study, and this effect should be taken into
account when designing samples of ecological networks in the future. We
see a similar qualitative pattern in empirical networks
(fig.~\ref{fig:geometric}(d)) where the FNR drops off quickly as a
function of observation effort, mediated by total richness. The
practical consequence of the bottom row of fig.~\ref{fig:geometric} when
conducting an analysis is whether there are enough total number of
observed interactions (the x-axis) for the threshold FNR we deem
acceptable (the y-axis) is feasible. This raises two points: first,
empirical data on interactions are subject to the practical limitations
of funding and human-work hours, and therefore existing data tend to
fall on the order of hundreds or thousands observations of individuals
per site. Clear aggregation of data on sampling effort has proven
difficult to find and a meta-analysis of network data and sampling
effort seems both pertinent and necessary, in addition to the effects of
aggregation of interactions across taxonomic scales (Gauzens \emph{et
al.} 2013; Giacomuzzo \& Jordán 2021). This inherent limitation on
in-situ sampling means we should optimize where we sample across space
so that for a given number of samples, we obtain the maximum information
possible. Second, what is meant by ``acceptable'' FNR? This raises the
question: does a shifting FNR lead to rapid transitions in our ability
inference and predictions about the structure and dynamics of networks,
or does it produce a roughly linear decay in model efficacy? We explore
this in the final section.

We conclude this section by advocating for the use of neutral models
similar to above to generate expectations about the number of
false-negatives in a dataset of a given size. This could prove fruitful
both for designing surveys of interactions but also because we may want
to incorporate models of imperfect detection error into predictive
interactions models, as Joseph (2020) does for species occurrence
modeling. Additionally, we emphasize that one must consider the context
for sampling---is the goal to detect a particular species (as in
fig.~\ref{fig:geometric}(c)), or to get a representative sample of
interactions across the species pool? These arguments are
well-considered when sampling individual species (Willott 2001), but
have not yet been adopted for designing samples of communities.

@@

Here we show how to incorporate uncertainty into model predictions of
interaction probability to account for imperfect observation (both
false-negatives and false-positives). Models for interaction prediction
typically yield a probability of interaction between each pair of
species, \(p_{ij}\). When these are considered with uncertainty, it is
usually model-uncertainty, e.g.~the variance in the interaction
probability prediction across several cross-validation folds, where the
data is split into training and test sets several times. The method we
introduce adjusts the value of a model's predictions to produce a
distribution of interaction probabilities corrected by a given
false-negative-rate \(p_{fn}\) and false-positive-rate \(p_{fp}\)
(outlined in figure fig.~\ref{fig:resampling_concept}). First we
describe how to sample from this distribution of adjusted interaction
probabilities via simulation, and show that this distribution can be
well-approximated analytically.

@@

To get an estimate of each interaction probability that accounts for
observation error, we resample the output prediction from an arbitrary
model for each interaction \(p_{ij}\) by simulating a set of several
`particles.' Each particle is a realization of an interaction
\emph{actually occurring} assuming \(p_{ij}\) is a correct estimate of
the probability of an interaction being \emph{observed}. Each particle
starts as being drawn as true or false according to \(p_{ij}\), and then
adjusting this by the rate of observation error given by \(p_{fp}\) and
\(p_{fn}\) to yield a single boolean outcome for each particle
(``Resampling'' within fig.~\ref{fig:resampling_concept}(b)). Across of
many particles, the resulting frequency of `true' outcomes is a single
resample of the probability \(p_{ij}^*\) that the interaction actually
\emph{occurred}, not just that it was \emph{observed}. Across several
samples each of several particles, this forms a distribution of
probabilities which are adjusted by the true and false-negative-rates.

@@

To test this we use the dataset from Hadfield \emph{et al.} (2014) that
describes host-parasite interaction networks sampled across 51 sites,
and the same method as Strydom \emph{et al.} (2021) to extract latent
features for each species in this dataset based on applying PCA to the
co-occurrence matrix. We then predict a metaweb (equivalent to
predicting true or false for an interaction between each species pair,
effectively a binary classification problem) from these species-level
features using four candidate models for binary classification---three
often used machine-learning (ML) methods (Boosted Regression Tree (BRT),
Random Forest (RF), Decision Tree (DT)), and one naive model from
classic statistics (Logistic Regression (LR)). Each of the ML models are
bootstrap aggregated (or bagged) with 100 replicates each. We partition
the data into 80-20 training-test splits, and then seed the training
data with false negatives at varying rates, but crucially do
\emph{nothing} to the test data. We fit all of these models using
MLJ.jl, a high-level Julia framework for a wide-variety of ML models
(Blaom \emph{et al.} 2020). We evaluate the efficacy of these models
using two common measures of binary classifier performance: the area
under the receiver-operator curve (ROC-AUC) and the area under the
precision-recall curve (PR-AUC), for more details see Poisot (2022).
Here, PR-AUC is slightly more relevant as it is a better indicator of
prediction of false-negatives. The results of these simulations are
shown in fig.~\ref{fig:addedfnr}(a \& b).
